Bioverse: Origins of Life

### 2024-07-05 src/tex/ms.tex
Motivate semianalytical and bioverse approaches

\subsection{Semi-analytical approach}\label{sec:met-semianalytical}
First, we apply a Bayes Factor Design Analysis~\citep{Schoenbrodt2018} to assess the expected probabilities of obtaining true negative or true positive evidence for the hypothesis above, as well as the probability for misleading or inconclusive evidence.

@@

\subsection{Exoplanet survey simulations with Bioverse}
To assess the information gain of a real exoplanet survey, we employed Bioverse~\citep{Bixel2021}, a framework that integrates multiple components including statistically realistic simulations of exoplanet populations, a survey simulation module, and a Bayesian hypothesis testing module to evaluate the statistical power of different observational strategies.
The general approach is as follows:
\begin{enumerate}
\item \textbf{Exoplanet population synthesis:} We populate the Gaia Catalogue of Nearby Stars~\citep{Smart2021} with synthetic exoplanets whose orbital parameters and planetary properties reflect our current understanding of exoplanet demographics~\citep{Bergsten2022}.
    \item \textbf{Survey simulation:} We simulate the detection and characterization of these exoplanets with a hypothetical survey, taking into account the survey's sensitivity, target selection, and observational biases.
To model the sensitivity of the information gain of a proposed mission to sample selection and survey strategy, we conduct survey simulations with Bioverse using different sample sizes and survey strategies.
    \item \textbf{Hypothesis testing:} We employ Bayesian statistical methods to evaluate the likelihood that a given survey would detect a specified statistical trend in the exoplanet population and estimate the precision with which the survey could constrain the parameters of that trend.
    A common definition of the null hypothesis $H_0$, which is also applied here, is that there is no relationship between the independent variable (here: maximum NUV flux) and the dependent variable (here: biosignature occurrence).
    The alternative hypothesis $H_1$ proposes a specific relationship parameterized by $\theta$.
    If the dependent variable is of binary nature such as in the case of biosignature detections, we model the likelihood function as
    \begin{equation}
\mathcal{L}(y \mid \boldsymbol{\theta})=\prod_i^N\left[y_i h\left(\boldsymbol{\theta}, x_i\right)+\left(1-y_i\right)\left(1-h\left(\boldsymbol{\theta}, x_i\right)\right)\right],
\end{equation}
where here $y_i \in \{0,1\}$ is the biosignature detection variable, $x_i$ is the maximum NUV flux, and $h(\boldsymbol{\theta}, x_i)$ is the probability of detecting a biosignature given the maximum NUV flux and the model parameters $\boldsymbol{\theta}$.
We then use a nested sampling method~\citep{Speagle2020} to compute the Bayesian evidence for the null and alternative hypotheses and estimate the strength of evidence for the alternative hypothesis.

@@

\begin{figure}
    \script{inhabited_FGKM.py}
    \begin{centering}
        \includegraphics[width=\hsize]{figures/inhabited_FGKM}
        \caption{Host stars of all transiting \glspl{EEC} and inhabited planets.
        In the FGK sample, most \glspl{EEC} and all inhabited planets orbit K~dwarfs.
        In an M~dwarf sample of the same size, the fraction of inhabited planets is larger.
        }
        \label{fig:inhabited_FGKM}
    \end{centering}
\end{figure}

@@

\begin{figure*}
    \script{hz_nuv_evo.py}
    \begin{centering}
        \includegraphics[width=\hsize]{figures/hz_nuv_evo}
        \caption{Interpolated stellar luminosity evolution (left) and evolution of the \gls{NUV} flux in the \gls{HZ} (right) as a function of host star mass.
        The scatter plots show age and host star mass of the transiting planets in the synthetic FGK sample; crosses denote the estimated \gls{NUV} values in \citet{Richey-Yowell2019}. % with \gls{EEC} highlighted in green.
        A few example tracks are shown, where extended overlap of \gls{HZ} occupancy and high \gls{NUV} flux (green sections) fulfill our requirement for abiogenesis.
        Planet~1 is an \gls{EEC} that never receives sufficient \gls{NUV} flux for abiogenesis.
        Planet~2 and Planet~3 enter the \gls{HZ} at different times and receive sufficient \gls{NUV} flux for different durations until their respective host star evolves below the threshold.
        }
        \label{fig:hz_nuv_evo}
    \end{centering}
\end{figure*}

@@

\subsubsection{Sampling strategy for testing a predicted minimum \gls{NUV} flux}
We show in Sect.~\ref{sec:results} that the constraining power for testing the hypothesis of a minimum past \gls{NUV} flux required for abiogenesis is sensitive on the occurrence of life, the value of this threshold flux, the sample size, and the distribution of sampled past \gls{NUV} fluxes.
In particular the last parameter offers an opportunity to optimize the survey strategy: Although constraints on a planet's UV history have generally large uncertainties~\citep[e.g.,][]{Richey-Yowell2023}, the likely maximum flux of a planet can be inferred and used as a proxy.
Sampling more extreme (low and high) values for the maximum flux increases the probability of obtaining true strong evidence for the hypothesis.
For large sample sizes $\gtrsim 200$, this strategy can push this probability into the \SI{90}{\percent} range.
\todo[inline]{and what about the probability of getting true strong evidence for H0?}

>>>>

\subsection{Semi-analytical approach}\label{sec:met-semianalytical}
First, we apply a Bayes Factor Design Analysis~\citep{Schoenbrodt2018} to assess the expected probabilities of obtaining true negative or true positive evidence for the hypothesis above, as well as the probability for misleading or inconclusive evidence, under idealized conditions.
This serves as a first-order estimate of the information content of a survey, before we take into account impacts from exoplanet demographics, sample selection, and survey strategy.

@@

\subsection{Exoplanet survey simulations with Bioverse}
A real exoplanet survey will be subject to observational biases, sample selection effects, and the underlying demographics of the planet sample.
To assess the information gain of a realistic exoplanet survey, we employed Bioverse~\citep{Bixel2021}, a framework that integrates multiple components including statistically realistic simulations of exoplanet populations, a survey simulation module, and a Bayesian hypothesis testing module to evaluate the statistical power of different observational strategies.
The general approach is as follows:
\begin{enumerate}
\item \textbf{Exoplanet population synthesis:} We populate the Gaia Catalogue of Nearby Stars~\citep{Smart2021} with synthetic exoplanets whose orbital parameters and planetary properties reflect our current understanding of exoplanet demographics~\citep{Bergsten2022}.
    \item \textbf{Survey simulation:} We simulate the detection and characterization of these exoplanets with a hypothetical survey, taking into account the survey's sensitivity, target selection, and observational biases.
To model the sensitivity of the information gain of a proposed mission to sample selection and survey strategy, we conduct survey simulations with Bioverse using different sample sizes and survey strategies.
    \item \textbf{Hypothesis testing:} We employ Bayesian statistical methods to evaluate the likelihood that a given survey would detect a specified statistical trend in the exoplanet population and estimate the precision with which the survey could constrain the parameters of that trend.
    A common definition of the null hypothesis $H_0$, which is also applied here, is that there is no relationship between the independent variable (here: maximum NUV flux) and the dependent variable (here: biosignature occurrence).
    The alternative hypothesis $H_1$ proposes a specific relationship parameterized by $\theta$.
    If the dependent variable is of binary nature such as in the case of biosignature detections, we model the likelihood function as
    \begin{equation}
\mathcal{L}(y \mid \boldsymbol{\theta})=\prod_i^N\left[y_i h\left(\boldsymbol{\theta}, x_i\right)+\left(1-y_i\right)\left(1-h\left(\boldsymbol{\theta}, x_i\right)\right)\right],
\end{equation}
where here $y_i \in \{0,1\}$ is the biosignature detection variable, $x_i$ is the maximum NUV flux, and $h(\boldsymbol{\theta}, x_i)$ is the probability of detecting a biosignature given the maximum NUV flux and the model parameters $\boldsymbol{\theta}$.
We then use a nested sampling method~\citep{Speagle2020} to compute the Bayesian evidence for the null and alternative hypotheses and estimate the strength of evidence for the alternative hypothesis.

@@

\begin{figure}
    \script{inhabited_FGKM.py}
    \begin{centering}
        \includegraphics[width=\hsize]{figures/inhabited_FGKM}
        \caption{Host stars of all transiting \glspl{EEC} and inhabited planets in a simulated transit survey.
        In the FGK sample, most \glspl{EEC} and all inhabited planets orbit K~dwarfs.
        In an M~dwarf sample of the same size, the fraction of inhabited planets is larger.
        }
        \label{fig:inhabited_FGKM}
    \end{centering}
\end{figure}

@@

\begin{figure*}
    \script{hz_nuv_evo.py}
    \begin{centering}
        \includegraphics[width=\hsize]{figures/hz_nuv_evo}
        \caption{Interpolated stellar luminosity evolution (left) and evolution of the \gls{NUV} flux in the \gls{HZ} (right) as a function of host star mass.
        The scatter plots show age and host star mass of the transiting planets in the synthetic FGK sample; crosses denote the estimated \gls{NUV} values in \citet{Richey-Yowell2019}. % with \gls{EEC} highlighted in green.
        A few example tracks for an example threshold flux of $F_\mathrm{NUV, min} = \var{NUV_thresh}\,\SI{}{\erg\per\second\per\centi\meter\squared}$ are shown; extended overlap of \gls{HZ} occupancy and high \gls{NUV} flux (green sections) fulfills our requirement for abiogenesis.
        Planet~1 is an \gls{EEC} that never receives sufficient \gls{NUV} flux for abiogenesis.
        Planet~2 and Planet~3 enter the \gls{HZ} at different times and receive sufficient \gls{NUV} flux for different durations until their respective host star evolves below the threshold.
        }
        \label{fig:hz_nuv_evo}
    \end{centering}
\end{figure*}

@@

\subsection{Sampling strategy for testing a predicted minimum \gls{NUV} flux}
We show in Sect.~\ref{sec:results} that the constraining power for testing the hypothesis of a minimum past \gls{NUV} flux required for abiogenesis is sensitive on the occurrence of life, the value of this threshold flux, the sample size, and the distribution of sampled past \gls{NUV} fluxes.
In particular the last parameter offers an opportunity to optimize the survey strategy: Although constraints on a planet's UV history have generally large uncertainties~\citep[e.g.,][]{Richey-Yowell2023}, the likely maximum flux of a planet can be inferred and used as a proxy.
Sampling more extreme (low and high) values for the maximum flux increases the probability of obtaining true strong evidence for the hypothesis.
For large sample sizes $\gtrsim 200$, this strategy can push this probability into the \SI{90}{\percent} range.
\todo[inline]{and what about the probability of getting true strong evidence for H0?}
